EGGs are adhesive!

### 2025-03-13 calco/mainTmp.tex
minima

\newcommand{\Ob}{\mathcal{O}b}
\newcommand{\Hom}{\mathcal{H}om}
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Reg}{\mathcal{Reg}}
\newcommand{\Mono}{\mathcal{Mono}}
\newcommand{\initial}{\mymathbb{0}}
\newcommand{\terminal}{\mathds{1}}
%\newcommand{\eg}[1]{\mathbf{EqGraph}_{\textbf {\textup{#1}}}}
\newcommand{\eg}[0]{\mathbf{EGG}}

@@

\noindent
\parbox{7.5cm}{\begin{restatable}{lemma}{mps}\label{prop:kerset}
	Suppose that in $\Set$ the cube in the diagram on the left is given, whose top face is a pushout, the left and bottom faces are pullbacks,  and $n\colon B\mto D$ is an injection. 
	Then it holds:
	\begin{enumerate}
		\item the right face of the cube is a pullback too;
		\item the right square, made by the kernel pairs of the vertical arrows, is a pushout.
	\end{enumerate}
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=5pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar@{>->}[rr]_{g} & & D }}

@@

\noindent
\begin{minipage}[l]{.73\linewidth}
\begin{definition}
An \emph{algebraic signature} $\Sigma$ is a pair $(O_\Sigma, \ari_\Sigma)$ given by a \emph{set of operations} $O_\Sigma$ and an \emph{arity function} $\ari_\Sigma\colon O_\Sigma \to \mathbb{N}$. 
%
We define the \emph{hypergraph $\mathcal{G}_\Sigma$ associated with $\Sigma$} taking $O_\Sigma$ as set of hyperedges, $\{1\}$ as set of nodes, so that $\{1\}^\star$ is $\mathbb{N}$, $\ari_\Sigma$ as the source function and $\gamma_1$ as target function, which always picks the element $1$. The category $\hyps$ of \emph{algebraically labelled hypergraphs} is the slice category $\hyp/\mathcal{G}^\Sigma$.
\end{definition}

@@

\begin{corollary}\label{cor:mono1}
	An arrow $h = (h_E, h_V, h_C): \mathcal{G \to H}$ in $\EqHyp$ is a mono if and only if $(h_E, h_V)$ is a mono in $\hyp$.
\end{corollary}

@@

\begin{definition}Let $\Sigma$ be an algebraic signature.
	An object $(\mathcal{H}, l)$ of $\EqHyp_\Sigma$ is a \emph{term graph with equivalence} if $l\colon T(\mathcal{H})\to \mathcal{G}^\Sigma$ is a term graph. We denote by $\EqTG_{\Sigma}$ the full subcategory of $\EqHyp_\Sigma$ so defined and by $J_{\Sigma}$ the corresponding inclusion functor.
 \end{definition}

\noindent 
\parbox{9.5cm}{\begin{remark}\label{rem:obv}
Let $T_\Sigma$ be the functor sending $(\mathcal{H}, l)$ to $(T(\mathcal{H}), T(l))$.   Notice that, by definition, $(\mathcal{H}, l)$ is in $\EqTG_{\Sigma}$ if and only if $T(\mathcal{H})$ is in $\tg$. Thus there exists a functor $S_\Sigma$ as in the diagram on the right.
\end{remark}}\hfill \parbox{4cm}{\xymatrix@R=15pt{\EqTG_{\Sigma} \ar[r]_{J_\Sigma}  \ar[d]_{S_\Sigma}& \EqHyp_\Sigma \ar[d]^{T_\Sigma} \\ \tg \ar[r]^{I_\Sigma}& \hyp_{\Sigma}}}

@@

 \begin{restatable}{proposition}{trm}\label{prop:term}
$\EqTG_{\Sigma}$ has equalizers, binary products and pullbacks and they are created by $J_\Sigma$.
 \end{restatable}

\noindent
\parbox{11cm}{\hspace{15pt}Let now $\mathcal{T}$ be the class of morphism in $\EqTG_{\Sigma}$ whose image through $J_\Sigma$ is in $\pbc_\Sigma$ and whose image through $S_\Sigma$ is a regular mono in $\tg$.  Unwrapping the definition and using \Cref{lem:reg}, we have that a morphism $(h_E, h_V, h_Q)\colon (\mathcal{G},l)\to (\mathcal{H}, l')$ is in $\mathcal{T}$ if and only if all of its components are injections and the square on the right is a pullback.} \hfill 
\parbox{2cm}{\xymatrix{V_{\mathcal{G}}\ar@{>->}[r]^{h_V} \ar@{>>}[d]_{q_{\mathcal{H}}}& V_{\mathcal{H}} \ar@{>>}[d]^{q_{\mathcal{H}}}\\ Q_{\mathcal{G}} \ar@{>->}[r]_{h_Q}& Q_{\mathcal{H}}} }

@@

\begin{restatable}{proposition}{po}\label{prop:po}
$\EqTG_{\Sigma}$ has all $\mathcal{T}$-pushouts, which are created by $J_{\Sigma}$.
\end{restatable}

@@

We can now turn to pushouts. We refer again the reader to \Cref{proof:epo} for the details.

\noindent
\parbox{11.5cm}{\begin{restatable}{lemma}{epo}\label{lem:epo}
	Suppose that the square on the right is a pushout in $\EqHyp$ with $m$ an arrow in $\pbc$. If $\mathcal{G}_1$, $\mathcal{G}_2$ and $\mathcal{G}_3$ are e-hypergraphs, then $\mathcal{P}$ is an object of $\egg$ too.
\end{restatable}}\hfill 
\parbox{4cm}{\xymatrix@R=15pt{\mathcal{G}_1 \ar[r]^{h}\ar[d]_{m}&\mathcal{G}_2\ar[d]^{n}\\\mathcal{G}_3\ar[r]_{z}&\mathcal{P}}}

Let now $\pbe$ be the class of arrows in $\egg$ which are sent to $\pbc$ by $I$. By definition and \Cref{lem:pbmono} such class is closed under composition, decomposition, contains all isomorphisms and it is stable under pullbacks. Using \Cref{lem:pbmono,lem:epo} we can also deduce that it is stable under pushouts. Then \Cref{thm:slice-functors,cor:eqade} yield the following.

\begin{corollary}\label{cor:pbe}
	$\egg$ is $\pbe$-adhesive.
\end{corollary}

\subsection{E-term graphs, a.k.a. EGGs}

We turn now to the labelled context.

\begin{definition} We say that an object $(\mathcal{H}, l)$ of $\EqHyp_\Sigma$ is a \emph{labelled e-hypergraph} if $\mathcal{H}$ is an e-hypergraph. We define the category $\egg_\Sigma$ as the full subcategory of $\EqHyp_\Sigma$ given by labelled e-hypergraphs. We will denote by $j_\Sigma$ the corresponding inclusion functor .
\end{definition}

To prove some adhesivity property of $\egg_\Sigma$ let us begin with the following elementary, but useful observation.

\begin{remark}\label{rem:algegg}
	Given a signature $\Sigma = (O_\Sigma, \ari_\Sigma)$,
	the hypergraph 
	$R(\mathcal{G}^\Sigma) = (O_\Sigma, 1, 1, \ari_\Sigma, \gamma_1, \id{1})$
	is an object of $\egg$.
	Indeed, under the identification of $1^\star$ with $\mathbb{N}$, the kernel $(S, \pi_1, \pi_2)$  of $\id{\mathbb{N}} \circ \ari_\Sigma$, is given by
	\[S:=\{(o_1, o_2)\in O_\Sigma \times O_\Sigma \mid \ari_\Sigma(o_1)=\ari_\Sigma(o_2)\}\]
	equipped with the two projections.  Onthe other hand $\id{\mathbb{N}} \circ \gamma_1 \circ \pi_1$ and $\id{\mathbb{N}}\circ \gamma_1 \circ \pi_2$ are both the function $O_\Sigma \to \mathbb{N}$ constant in $1$.
\end{remark}

\Cref{rem:algegg} now implies at once the following result.
\begin{proposition}\label{prop:varie}
	Let $\Sigma$ be an algebraic signature. Then the following hold true:
	\begin{enumerate}
		\item $\egg_\Sigma$ is equivalent to $\egg/R(\mathcal{G}^\Sigma)$.
		\item There exists a functor $W_\Sigma \colon \egg_\Sigma\to \Sigma$ forgetting the labeling which creates all colimits, pullbacks and equalizers. 
	\end{enumerate}
\end{proposition}

Let $\pbe_\Sigma$ be the class of morphisms in $\egg_\Sigma$ whose image in $\egg$ lies in $\pbe$. Notice that $\pbe_\Sigma$ is also the class of arrows whose image through $j_\Sigma$ is in $\pbc_\Sigma$.

By \Cref{thm:slice-functors} we get the following.

\begin{corollary}\label{cor:eggade1}
	$\egg_\Sigma$ is $\pbe_\Sigma$-adhesive.
\end{corollary}

We turn now to term graphs with equivalence.

\begin{definition}\label[definintion]{def:GEqTGs}
	Given a signature $\Sigma$,  we say that an object $(\mathcal{H}, l)$ of $\EqTG_{\Sigma}$ is a \emph{e-term graph} if $\mathcal{H}$ is an e-hypergraph. We define the category $\eg$ as the full subcategory of $\EqTG_{\Sigma}$ given by e-term graphs and denote by $k_\Sigma$ the corresponding inclusion.
\end{definition}

\begin{remark}
	By definition we also have an inclusion functor $K_\Sigma\colon \eg\to \egg_\Sigma$.
\end{remark}

\noindent 
\parbox{7.5cm}{\begin{remark}
	Now that we have introduced all the characters of this work, it is worth to give a visual map of all the categories that we have introduced in this paper and of the relationships between them. We will use the curved arrows to denote full and faithful inclusions.
\end{remark}}\hfill \parbox{4cm}{\xymatrix@R=15pt@C=15pt{\eg \ar@{^{(}->}[r]^-{K_\Sigma} \ar@{^{(}->}[d]_{k_\Sigma} & \egg_\Sigma \ar@{^{(}->}[d]^{j_\Sigma} \ar[r]^{W_\Sigma}& \egg \ar@{^{(}->}[d]^{I}\\ \EqTG_{\Sigma} \ar@{^{(}->}[r]_{J_\Sigma}\ar[d]_{S_\Sigma}& \EqHyp_\Sigma \ar[r]_{V_\Sigma} \ar[d]_{T_\Sigma}& \EqHyp \ar[d]_{T}\\ \tg \ar@{^{(}->}[r]_{I_\Sigma}& \hyp_{\Sigma} \ar[r]_{U_\Sigma}  &\hyp}}

We can now prove our last three results. The interested reader can find the proof in \Cref{proof:tlim}.

\begin{restatable}{proposition}{limt}\label{prop:limt}
	$\eg$ has equalizer, binary products and pullbacks and they are created by $k_\Sigma$.
\end{restatable}

Let now $\mathcal{T}_\Sigma$ be the class of morphisms of $\eg$ which are sent by $k_\Sigma$ to the class $\mathcal{T}$. Then \Cref{colim,lem:epo,prop:po} allow us to deduce the following result.

\begin{restatable}{proposition}{pot}\label{prop:pot}
	$\eg$ has $\mathcal{T}_\Sigma$-pushouts which are created by $k_\Sigma$.
\end{restatable}

Reasoning as in the previous sections, \Cref{prop:limt,prop:pot} now give us the following.

\begin{corollary}
	$\eg$ is $\mathcal{T}_\Sigma$-adhesive.
\end{corollary}

\commentato{ 
\begin{remark}\label[remark]{rem:sig_egg}
	Given a signature $\Sigma = (O_\Sigma, \ari_\Sigma)$,
	the hypergraph 
	$R(\mathcal{G}^\Sigma) = (O_\Sigma, 1, 1, \ari_\Sigma, \delta_1, \id{1})$
	is an object of $\egg$.
	Indeed, if $(S, \pi_1, \pi_2)$ is the kernel pair of $\id{1}^\star \circ \ari_\Sigma$,
	we have that $\id{1}^\star \circ \delta_1 \circ \pi_1$ and $\id{1}^\star \circ \delta_1 \circ \pi_2$ are both the contant functions of value $1$.
\end{remark}

@@

%\todo{Here we must introduce $\GEqTGs$}
>>>>>>> refs/remotes/origin/main}

@@

\begin{proof}\label{proof:kerset}
	We can start noticing that $m$, being the pullback of $n$ is mono too. This, in turn, entails that $m'$, being the pullback of $m$, is mono, and so $n'$ is injective too because in $\Set$, as in any adhesive category, monomorphisms are stable under pushouts.

	\noindent 
	\parbox{11.5cm}{\hspace{15pt}
	Let $\iota' \colon E'\mto C'$ and $\iota\colon E\to C$ be the inclusions of the complement of the images of of $m'$ and $m$ respectively . By \Cref{lem:po_set}, $(D', \{n', g'\circ \iota\})$ is a coproduct. We can also notice that $c\circ \iota'$ factors through $E$, as shown by the square on the right.}\hfill \parbox{4cm}{\vspace{-5pt}\xymatrix@R=15pt{E' \ar@{.>}[d]_{w} \ar@{>->}[r]_{\iota'} & C' \ar[d]^{c}\\ E \ar@{>->}[r]^{\iota} & C}}

@@

		By hypothesis the bottom face of the given cube is a pullback, therefore there exists an arrow $u\colon T_0\to A$ such that 
		\[f\circ u=b\circ t_2\circ l_0 \qquad m\circ u=c\circ \iota'\circ h_0\]

@@

\begin{proof}\label{proof:pbmono}
	Closure under composition and decomposition follows from \Cref{lem:pb1}, while stability follows from the first point of \Cref{prop:kerset} because pushouts along injections are also pullbacks by \Cref{prop:regmono}. We are left with stability under pullbacks. 
	Let $h\colon \mathcal{G}_1\to \mathcal{G}_3$ be an arrow in $\pbc$ and $k\colon \mathcal{G}_2\to \mathcal{G}_3$ be another morphism of $\EqHyp$, with $\mathcal{G}_{i}=(E_i, V_i, Q_i, s_i, t_i, q_i)$.

@@

\po*
\begin{proof}\label{proof:tade}
	Suppose that the square on the left below is a pushout in $\EqHyp_\Sigma$, with $h$ in $\mathcal{T}$. Then, by \Cref{rem:t} the square on the right is a pushout and by the definition of $\mathcal{T}$ and \Cref{rem:obv}, $T_\Sigma(h)$ is a regular mono in $\tg$.
	\[\xymatrix{J_\Sigma(\mathcal{G}_0, l_0) \ar[r]^{k} \ar[d]_{h}& J_\Sigma(\mathcal{G}_1, l_1) \ar[d]^{p} &T_\Sigma (J_\Sigma(\mathcal{G}_0, l_0)) \ar[r]^{T_{\Sigma}(k)} \ar[d]_{T_{\Sigma}(h)}& T_\Sigma (J_\Sigma(\mathcal{G}_1, l_1)) \ar[d]^{T_\Sigma(p)}\\
J_\Sigma(\mathcal{G}_1, l_1) \ar[r]_{q}& J_\Sigma(\mathcal{H}, l) & T_\Sigma(J_\Sigma(\mathcal{G}_1, l_1)) \ar[r]_{T_\Sigma(q)}& T_\Sigma(J_\Sigma(\mathcal{H}, l))}\]

@@

	\noindent 
	\parbox{9.5cm}{ \vspace{-.8cm}\hspace{15pt}
	Thus for each $d$ in $\D$, there exists an arrow $a_d:U\to U_i$ making the diagram on the right commutative.
Then we have
\begin{align*}
	&l_d^\star\circ m^\star \circ q^\star \circ t \circ u_1    = q_d^\star\circ (\pi_V^d)^\star \circ t \circ u_1 \\
	= &q_d^\star \circ t_d \circ \pi_E^d \circ u_1
	= q_d^\star \circ t_d \circ u_1^d \circ a_d 
	= q_d^\star \circ t_d \circ u_2^d \circ a_d\\ 
	=& q_d^\star \circ t_d \circ \pi_E^d \circ u_2 = q_d^\star \circ (\pi_V^d)^\star \circ t \circ u_2 = l_d^\star \circ m^\star \circ q^\star \circ t \circ u_2
\end{align*}}\hfill \parbox{4cm}{\vspace{-1cm}	\xymatrix@R=20pt{
		U \ar[r]^{u_1} \ar[d]_{u_2} \ar@{.>}[dr]_{a_d} & A \ar[dr]^{\pi_E^d} & \\
		A \ar[dr]_{\pi_E^d} & U_d \ar[r]^{u_1^d} \ar[d]_{u_2^d}& A_d \ar[d]^{q_d^\star \circ s_d} \\
		& A_i  \ar[r]_{q_d^\star \circ s_d}& Q_d^\star
	}}

@@

\begin{remark}\label{rem:kleene}
	Let $\delta_1\colon 1\to \mathbb{N}$ be the arrow which picks $1\in \mathbb{N}$. Consider the

	\noindent
	\parbox{11.5cm}{ inclusion $v_1\colon X\to X^\star $  defined by the first point of \Cref{prop:fact}. By extensivity we know that the square on the right is a pullback.}\hfill \parbox{4cm}{\vspace{-.5cm}\xymatrix@R=15pt{X \ar[r]_{!_{X}} \ar[d]_{!_{X}}& 1\ar[d]^{\delta_1} \\1 \ar[r]^{\lgh_{X}} & \mathbb{N}} }

	\noindent 
	\parbox{4cm}{\vspace{-.2cm} \xymatrix@C=15pt@R=1pt{&X\ar[dd]|\hole_(.65){!_{X}}\ar@{->}[rr]^{!_X} \ar@{>}[dl]_{f} && 1 \ar[dd]^{\delta_1} \ar[dl]^{\id{1}} \\ Y  \ar[dd]_{!_Y}\ar@{>}[rr]^(.7){!_Y} & & 1 \ar[dd]_(.3){\delta_1}\\&X^\star\ar@{>->}[rr]|(.53)\hole^(.70){\lgh_{X}} \ar@{>->}[dl]_{f^\star} && \mathbb{N} \ar[dl]^{\id{N}} \\Y^\star \ar@{>->}[rr]_{\lgh_{Y}} & & \mathbb{N}}}\hfill \parbox{9.5cm}{\vspace{-.3cm}\hspace{15pt}Let now $f\colon X\to Y$ be an arrow. Then we can build the cube aside which, by what we have just observed above, has pullbacks as back, front and right faces. Thus, by \Cref{lem:pb1} its left face is a pullback too.}

\end{remark}

>>>>

\newcommand{\Ob}{\mathcal{O}b}
\newcommand{\Hom}{\mathcal{H}om}
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Reg}{\mathcal{Reg}}
\newcommand{\Mono}{\mathcal{Mono}}
\newcommand{\initial}{\mymathbb{0}}
\newcommand{\terminal}{\mathds{1}}
\newcommand{\eg}[1]{\mathbf{EqGraph}_{\textbf {\textup{#1}}}}
%\newcommand{\egg}[1]{\mathbf{EGG}_{\textbf {\textup{#1}}}}

@@

\noindent
\parbox{7.5cm}{\begin{restatable}{lemma}{mps}\label{prop:kerset}
	Suppose that in $\Set$ the cube in the diagram on the left is given, whose top and bottom faces are pushout, the left face a pulllback and $m\colon A\mto C$ is an injection. 
	Then it holds
	\begin{enumerate}
		\item the right face of the cube is a pullback too;
		\item the right square, made by the kernel pairs of the vertical arrows, is a pushout.
	\end{enumerate}
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=5pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar@{>->}[rr]_{g} & & D }}

@@

\noindent
\begin{minipage}[l]{.73\linewidth}
\begin{definition}
An \emph{algebraic signature} $\Sigma$ is a pair $(O_\Sigma, \ari_\Sigma)$ given by a \emph{set of operations} $O_\Sigma$ and an \emph{arity function} $\ari_\Sigma\colon O_\Sigma \to \mathbb{N}$. 
%
We define the \emph{hypergraph $\mathcal{G}_\Sigma$ associated with $\Sigma$} taking $O_\Sigma$ as set of hyperedges, $\{1\}$ as set of nodes, so that $\{1\}^\star$ is $\mathbb{N}$, $\ari_\Sigma$ as the source function and $\delta_1$ as target function, which always picks the element $1$. The category $\hyps$ of \emph{algebraically labelled hypergraphs} is the slice category $\hyp/\mathcal{G}^\Sigma$.
\end{definition}

@@

\begin{corollary}\label{cor:mono1}
	An arrow $(h_E, h_V, h_C): \mathcal{G \to H}$ in $\EqHyp$ is a mono if and only if $(h_E, h_V)$ is a mono in $\hyp$.
\end{corollary}

@@

\begin{definition}Let $\Sigma$ be an algebraic signature.
	An object $(\mathcal{H}, l)$ of $\EqHyp_\Sigma$ is a \emph{term graph with equivalence} if $l\colon T(\mathcal{H})\to \mathcal{G}^\Sigma$ is a term graph. We denote by $\EqTG_{\Sigma}$ the full subcategory of $\EqHyp_\Sigma$ so defined and by $I_{\eq}$ the corresponding inclusion functor.
 \end{definition}

\noindent 
\parbox{9.5cm}{\begin{remark}\label{rem:obv}
Let $T_\Sigma:  \EqHyp_{\Sigma} \to \hyp_{\Sigma}$ the forgetful functor lifting
$T:  \EqHyp \to \hyp$.
%
%$T_\Sigma = T\circ V_\Sigma$.   
Notice that, by definition, $(\mathcal{H}, l)$ is in $\EqTG_{\Sigma}$ amounts to say that $(T(\mathcal{H}), l)$ is in $\tg$. 
Thus there exists a functor $S_\Sigma$ as in the diagram on the right.
\end{remark}}\hfill \parbox{4cm}{\xymatrix@R=15pt{\EqTG_{\Sigma} \ar[r]_{I_\eq}  \ar[d]_{S_\Sigma}& \EqHyp_\Sigma \ar[d]^{T_\Sigma} \\ \tg \ar[r]^{I_\Sigma}& \hyp_{\Sigma}}}

@@

 \begin{restatable}{proposition}{trm}\label{prop:term}
$\EqTG_{\Sigma}$ has equalizers, binary products and pullbacks and they are created by $I_\eq$.
 \end{restatable}

\noindent
\parbox{11cm}{\hspace{15pt}Let now $\mathcal{T}$ be the class of morphism in $\EqTG_{\Sigma}$ whose image through $I_\eq$ is in $\pbc_\Sigma$ and whose image through $S_\Sigma$ is a regular mono in $\tg$.  Unwrapping the definition and using \Cref{lem:reg}, we have that a morphism $(h_E, h_V, h_Q)\colon (\mathcal{G},l)\to (\mathcal{H}, l')$ is in $\mathcal{T}$ if and only if all of its components are injections and the square on the right is a pullback.} \hfill 
\parbox{2cm}{\xymatrix{V_{\mathcal{G}}\ar@{>->}[r]^{h_V} \ar@{>>}[d]_{q_{\mathcal{H}}}& V_{\mathcal{H}} \ar@{>>}[d]^{q_{\mathcal{H}}}\\ Q_{\mathcal{G}} \ar@{>->}[r]_{h_Q}& Q_{\mathcal{H}}} }

@@

\begin{restatable}{proposition}{po}\label{prop:po}
$\EqTG_{\Sigma}$ has all $\mathcal{T}$-pushouts, which are created by $I_{\eq}$.
\end{restatable}

@@

We can now turn to pushouts.

\begin{lemma}\label{lem:epo}
	Suppose that the square on the right is a pushout in $\EqHyp$ with $m$ a regular mono. If $\mathcal{G}_1$, $\mathcal{G}_2$ and $\mathcal{G}_3$ are e-hypergraphs, then $\mathcal{P}$ is an e-hypergraph too.
	following pushout square in $\EqHyp$.
	\[\xymatrix{\mathcal{G}_1 \ar[r]^{h}\ar[d]_{m}&\mathcal{G}_2\ar[d]^{n}\\\mathcal{G}_3\ar[r]_{k}&\mathcal{P}}\]
	with $m$ regular mono. 
\end{lemma}

\begin{proof}
	Let $\mathcal{P} = (A, B, C, s, t, q)$, $(K_i, \pi_i^1, \pi_i^2)$ the kernel pair of $q_i^\star \circ s_i$, and let $(U, u_1, u_2)$ the kernel pair of $q^\star \circ s$ .
	Consider then the following situation.
	\[\xymatrix@C=10pt@R=6pt{&A_1\ar[dd]|\hole_(.65){q_1^\star \circ s_1}\ar[rr]^{h_E} \ar[dl]_{m_E} && A_2 \ar[dd]^{q_2^\star \circ s_2} \ar[dl]_{k_E} \\ A_3 \ar[dd]_{q_3^\star \circ s_3}\ar[rr]^(.7){k_E} & & A \ar[dd]_(.3){q^\star \circ s}\\&{C_1^\star}\ar[rr]|\hole^(.65){h_C^\star}\ar[dl]^{m_C^\star} && {C_2^\star} \ar[dl]^{n_C^\star} \\{C_3^\star} \ar[rr]_{k_C^\star} & & C^\star}\]
	Since $m$ is regular mono, $m_E$ is mono ({\color{red}{inserire citazione}}).
	Then, by construction, the top face is a pushout, and since $\Set$ is adhesive, by \Cref{lem:mpo} , the square below is a pushout.
	\[\xymatrix{K_1\ar[r]^{f_k}\ar[d]_{f_m}&K_2\ar[d]^{f_n}\\K_3\ar[r]_{f_k}&U}\]
	Computing, we have
	\[
	\begin{split}
		q^\star \circ t \circ u_1 \circ f_n &= q^\star \circ t \circ n_E \circ \pi_2^1 \\ &= n_C^\star \circ q_2^\star \circ s_2 \circ \pi_2^1 \\&= n_C^\star \circ q_2^\star \circ s_2 \circ \pi_2^2 \\&=q^\star \circ t \circ u_2 \circ f_n 
	\end{split}
	\qquad
	\begin{split} q^\star \circ t \circ u_1 \circ f_k &= q^\star \circ t \circ k_E \circ \pi_3^1 \\ &= k_C^\star \circ q_3^\star \circ s_3 \circ \pi_3^1 \\&= k_C^\star \circ q_3^\star \circ s_3 \circ \pi_3^2 \\&=q^\star \circ t \circ u_2 \circ f_k
	\end{split}
	\]
	By universal property of pushouts, we deduce $q^\star \circ t \circ u_1 = q^\star \circ t \circ u_2$, and the thesis follows.
\end{proof}

\begin{corollary}\label{cor:rege}
		An arrow $h: \mathcal{G \to H}$  is a regular mono in $\egg$ if and only if $I(h)$ is a regular mono in $\EqHyp$. 
\end{corollary}

Let now $\pbe$ be the class of all regular monos in $\egg$ which are sent to $\pbc$ by $I$. By definition and \Cref{lem:pbmono} such class is closed under composition, decomposition, contains all isomorphisms and it is stable under pullbacks. Using \Cref{cor:rege} we can also deduce that it is stable under pushouts. Then \Cref{thm:slice-functors,cor:eqade} yield the following.

\begin{corollary}
	$\egg$ is $\pbe$-adhesive.
\end{corollary}

\subsection{E-Term graphs, a.k.a. EGGs}

\begin{remark}\label[remark]{rem:sig_egg}
	Given a signature $\Sigma = (O_\Sigma, \ari_\Sigma)$,
	the hypergraph 
	$R(\mathcal{G}^\Sigma) = (O_\Sigma, 1, 1, \ari_\Sigma, \delta_1, \id{1})$
	is an object of $\egg$.
	Indeed, if $(S, \pi_1, \pi_2)$ is the kernel pair of $\id{1}^\star \circ \ari_\Sigma$,
	we have that $\id{1}^\star \circ \delta_1 \circ \pi_1$ and $\id{1}^\star \circ \delta_1 \circ \pi_2$ are both the contant functions of value $1$.
\end{remark}

@@

%\todo{Here we must introduce $\GEqTGs$}

@@

\begin{proof}\label{proof:kerset}
	We can start noticing that $m'$, being the pullback of $m$ is mono too. This, in turn, entails that $n$ and $n'$ are mono too because in $\Set$, as in any adhesive category, monomorphisms are stable under pushouts.

	\noindent 
	\parbox{11.5cm}{\hspace{15pt}
	Let $\iota' \colon E'\mto C'$ and $\iota\colon E\to C$ be the inclusions of the complement of the images of of $m'$ and $m$ respectively . By \Cref{lem:po_set}, $(D', \{n', g'\circ \iota\})$ and $(D, \{n, g\circ \iota\})$ are coproducts. We can also notice that $c\circ \iota'$ factors through $E$, as shown by the square on the right.}\hfill \parbox{4cm}{\vspace{-5pt}\xymatrix@R=15pt{E' \ar@{.>}[d]_{w} \ar@{>->}[r]_{\iota'} & C' \ar[d]^{c}\\ E \ar@{>->}[r]^{\iota} & C}}

@@

		By adhesivity of $\Set$ and \Cref{prop:regmono} the bottom face of the given cube is a pullback, therefore there exists an arrow $u\colon T_0\to A$ such that 
		\[f\circ u=b\circ t_2\circ l_0 \qquad m\circ u=c\circ \iota'\circ h_0\]

@@

\begin{proof}\label{proof:pbmono}
	Closure under composition and decomposition follows from \Cref{lem:pb1}, while stability under pushouts is the first point of \Cref{prop:kerset}. We are left with stability under pullbacks. 
	Let $h\colon \mathcal{G}_1\to \mathcal{G}_3$ be an arrow in $\pbc$ and $k\colon \mathcal{G}_2\to \mathcal{G}_3$ be another morphism of $\EqHyp$, with $\mathcal{G}_{i}=(E_i, V_i, Q_i, s_i, t_i, q_i)$.

@@

\po*
\begin{proof}\label{proof:tade}
	Suppose that the square on the left below is a pushout in $\EqHyp_\Sigma$, with $h$ in $\mathcal{T}$. Then, by \Cref{rem:t} the square on the right is a pushout and by the definition of $\mathcal{T}$ and \Cref{rem:obv}, $T_\Sigma(h)$ is a regular mono in $\tg$.
	\[\xymatrix{I_\eq(\mathcal{G}_0, l_0) \ar[r]^{k} \ar[d]_{h}& I_\eq(\mathcal{G}_1, l_1) \ar[d]^{p} &T_\Sigma (I_\eq(\mathcal{G}_0, l_0)) \ar[r]^{T_{\Sigma}(k)} \ar[d]_{T_{\Sigma}(h)}& T_\Sigma (I_\eq(\mathcal{G}_1, l_1)) \ar[d]^{T_\Sigma(p)}\\
I_\eq(\mathcal{G}_1, l_1) \ar[r]_{q}& I_\eq(\mathcal{H}, l) & T_\Sigma(I_\eq(\mathcal{G}_1, l_1)) \ar[r]_{T_\Sigma(q)}& T_\Sigma(I_\eq(\mathcal{H}, l))}\]

@@

	\noindent 
	\parbox{9.5cm}{ \hspace{15pt}
	Thus for each $d$ in $\D$, there exists an arrow $a_d:U\to U_i$ making the diagram on the right commutative.
Then we have
\begin{align*}
	&l_d^\star\circ m^\star \circ q^\star \circ t \circ u_1    = q_d^\star\circ (\pi_V^d)^\star \circ t \circ u_1 \\
	= &q_d^\star \circ t_d \circ \pi_E^d \circ u_1
	= q_d^\star \circ t_d \circ u_1^d \circ a_d 
	= q_d^\star \circ t_d \circ u_2^d \circ a_d\\ 
	=& q_d^\star \circ t_d \circ \pi_E^d \circ u_2 = q_d^\star \circ (\pi_V^d)^\star \circ t \circ u_2 = l_d^\star \circ m^\star \circ q^\star \circ t \circ u_2
\end{align*}}\hfill \parbox{4cm}{	\xymatrix{
		U \ar[r]^{u_1} \ar[d]_{u_2} \ar@{.>}[dr]_{a_d} & A \ar[dr]^{\pi_E^d} & \\
		A \ar[dr]_{\pi_E^d} & U_d \ar[r]^{u_1^d} \ar[d]_{u_2^d}& A_d \ar[d]^{q_d^\star \circ s_d} \\
		& A_i  \ar[r]_{q_d^\star \circ s_d}& Q_d^\star
	}}

@@

\section{Some properties of comma categories}
In this section we will briefly recall the definition of the comma category \cite{mac2013categories} associated to two functors and some of its properties.
\begin{definition}\index{category!comma -}
	Let $L\colon \A\to \X$ and  $R\colon \B\rightarrow \X$ be two functors with the same codomain, the \emph{comma category} $\comma{L}{R}$ is the category in which
	\begin{itemize}
		\item objects are triples $(A, B, f)$ with $A\in \A$, $B\in \B$, and $f\colon L(A)\rightarrow R(B)$; 
		\item a morphism $(A, B, f)\rightarrow (A', B', g)$ is a pair $(h, k)$ with $h\colon A\rightarrow A'$ in $\A$ and $k\colon B\rightarrow B'$ in $\B$ such that the following diagram commutes
		\[\xymatrix@C=25pt{L(A) \ar[r]^{L(h)} \ar[d]_{f} & L(A') \ar[d]^{g}\\ R(B) \ar[r]_{R(k)}& R(B')}\]
	\end{itemize}
\end{definition} 
We have two forgetful functors 	$U_L\colon \comma{L}{R}\to \A$ and $U_R\colon \comma{L}{R} \to \B$ given, respectively by
\[
\begin{split}
	\functor[l]{(A,B, f)}{(h,k)}{(A', B', g)}
	& \functormapsto
	\rfunctor{A}{h}{A'}
\end{split}\quad 
\begin{split}
	\functor[l]{(A,B, f)}{(h,k)}{(A', B', g)}
	& \functormapsto
	\rfunctor{B}{k}{B'}
\end{split}
\]
